Detecting Facial Expressions in Professional Tennis Matches

### 2017-01-21 Report06-12.Rmd
Report Alterations
referencing in bookdown form

---
title: "Detecting Facial Expressions in Professional Tennis Matches"
author: "Stephanie Kobakian, Mitchell O'Hara-Wild, Dianne Cook, Stephanie Kovalcik"
date: '2017'
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    fig_crop: yes
    toc: yes
    toc_depth: 2
  html_notebook: null
fontsize: 11pt
documentclass: article
bibliography: references.bib
---

```{r cache=FALSE, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
#packages:
library(pander)
library(ggplot2)
library(imager)
library(devtools)
library(knitcitations)
library(RefManageR)
library(readr)
library(knitr)
library(grid)
library(tidyverse)
library(magick)
library(kfigr)
library(UpSetR)

@@

#csv files:
ALLmetaIMG<-read_csv("ALLmetaIMG.csv", col_types = cols(type = col_factor(levels = c("Manual", "Animetrics", "Google", "Microsoft", "Skybiometry"))))
SceneAttributes<-read_csv("TableCSVs/SceneAttributes.csv")
FaceAttributes<-read_csv("TableCSVs/FaceAttributes.csv")
SolutionSpecs<-read_csv("TableCSVs/SolutionSpecifications.csv")

@@

We were able to try the online demos to see whether these software were viable, we used the following Image displayed below in `r figr("Trial-Image", TRUE, type="Figure")`:

```{r Trial-Image, caption="This is a player", message=TRUE, warning=TRUE, anchor="figure"}
imagesList0<-as.list("2016_HSA_R01_BTomic_AUS_vs_DIstomin_UZB_MS157_clip.0013.png")
overlayGgplot(imagesList0, ALLmetaIMGFaces, matchBox = FALSE)
```

@@

```{r Scene-Attributes-Table, anchor="table", caption="Attributes of a Scene of an image. The most appropriate option from the list was selected for each attribute."}
kable(SceneAttributes, format = "pandoc", booktabs = TRUE)
```

The aim was to collect specific information on each different face within the scene. To determine which of the sometimes many faces in the scene it would be reasonable for software to detect a standard was created for reasonable detection. This was based on the attributes provided in `r figr("Solution-Specifications-Table", TRUE, type="Table")`, below:

```{r Solution-Specifications-Table, results='asis', anchor="table"}
kable(SolutionSpecs, format = "pandoc", booktabs = TRUE, caption = "Software solutions specifications.")

@@

```{r Face-Attribute-Tables, results='asis', anchor="table"}
kable(FaceAttributes, format = "pandoc", booktabs = TRUE, caption = "Attributes of an individual face within an image. The most appropriate option from the list was selected for each attribute.")
```

@@

The program we created was an interactive application that worked by presenting an image, using the imager @imager package, from the sample of images that had not yet been considered, appearing underneath were a set of radio buttons corresponding to the Scene Attributes in `r figr("Scene-Attributes-Table", TRUE, type="Table")` as shown above.

@@

Using the data set^[ALLmetaIMG] of the combined API and manual results, we were able to compare the performance of the software.
Firstly, we considered how many individual faces the software were able to detect in `r figr("Face-Bounding-Box-Data", TRUE, type="Figure")` .

However, individual faces are not beneficial if they do not correspond to the faces manually annotated. `r figr("Face-Bounding-Box-UpSet", TRUE, type="Figure")` was created by defining groups depending on the softwares that recognized each particular face. The UpSetR [-@UpSetR] package helps visualise set intersections. Where in this circumstance Face Bounding Boxes may overlap on the same face, each bar shows the number of individual faces that have Face Bounding Boxes resulting from each of the softwares highlighted below the bar.

Following, we considered the sizes of the Boxes created in `r figr("Face-Bounding-Box-Data-Box", TRUE, type="Figure")`, this was an attempt to determine how the software performed when presented with smaller faces. The boxplot uses the size of the Face Bounding Boxes, and contrasts based on the software the Boxes were found from.

To determine how each software reacted to the same face we looked at an image that contained only

finding faces that were recognized and grouping them by the softwares

Skybiometry results accounted for the 255 smallest recognized faces. However this is not necessarily a benefit to this research, as these are not all faces.

@@

```{r}
ALLmetaIMGFaces
#for each face there
```

@@

```{r Face-Bounding-Box-Data, results='hide', type="figure", caption="Face Bounding Box Data:", anchor="figure"}
ggplot(ALLmetaIMG, aes(x = type, fill=type)) + geom_bar(position="dodge") +xlab("Facial Recognition Softwares")+scale_x_discrete(limits=c("Manual","Google","Microsoft","Skybiometry","Animetrics")) +ylab("Number of Potential Faces") + guides(fill=FALSE)
```

The bar chart in 
`r figr("Face-Bounding-Box-Data", TRUE, type="Figure")` above shows the number of Bounding Boxes produced by each software, comparing the height of the bars indicates that Google's Facial Recognition software recognized almost 1000 more faces than the next best software, Microsoft.

-Google more than others

```{r Face-Bounding-Box-UpSet, results='hide', type="figure", caption="Face Bounding Box Data:", anchor="figure"}
createUpSet(ALLmetaIMGFaces)
```

`r figr("Face-Bounding-Box-Upset", TRUE, type="Figure")` Bar Chart shows the Face Bounding Boxes that were recognized by multiple software.
The largest group with 809 faces is Face Bounding Boxes only found by Manual annotations. The following group was Boxes representing the same faces being recognized by both Manual annotations and the Google API.

On average, Google maps the largest boxes around faces, and has the largest box recognized in the set. On average, the smallest Face Boxes are created by Skybiometry.

```{r I3, results=TRUE, caption="2016_SC2_R01_ATomljanovic_AUS_vs_KBondarenko_UKR_WS1112_clip.0053.png", anchor="figure"}
imagesList3<-as.list("2016_SC2_R01_ATomljanovic_AUS_vs_KBondarenko_UKR_WS1112_clip.0053.png") 
overlayGgplot(imagesList3, ALLmetaIMGFaces, matchBox = FALSE)
```

As before, the 'face' is not as we would hope it would be classified. Just right of the center, the smaller box actually captures a fist, not a face.

```{r I4,results=TRUE, caption="2016_RLA_R04_NDjokovic_SRB_vs_GSimon_FRA_MS401_clip.0003.png", anchor="figure"}
imagesList4<-as.list("2016_RLA_R04_NDjokovic_SRB_vs_GSimon_FRA_MS401_clip.0003.png")
overlayGgplot(imagesList4, ALLmetaIMG, matchBox = FALSE)
```

I4 shows the sensitivity of the software. We may consider this too sensitive as it recognized the eye and nose of the ball boy on the court.

```{r I5, results=TRUE, caption="2016_SC2_R01_AKerber_GER_vs_MDoi_JPN_WS1332_clip.0083.png", anchor="figure"}
imagesList5<-("2016_SC2_R01_AKerber_GER_vs_MDoi_JPN_WS1332_clip.0083.png") 
overlayGgplot(imagesList5, ALLmetaIMG, matchBox = FALSE)
```

As seen in this image of the crowd, the software is performing reasonably well. It does recognize the faces, yet only one not recognized by the other software, this may be due to the sunglasses.

This image demonstrates that the it is not necessarily the smaller faces that Skybiometry recognizes, but it puts smaller bounds on the facial features than other software.

```{r Type-Matches-Manual, anchor="table"}
library(descr, quietly = TRUE)
NoManual<-ALLmetaIMG %>% filter(ALLmetaIMG$type!="Manual")
NoManual<-NoManual %>% droplevels(NoManual$type)
Accuracy<-with(NoManual, table(NoManual$matchesManual, NoManual$type))
kable((Accuracy))
```

@@

@@

Table 4 shows whether the potential Face Bounding Boxes match faces that were annotated during the manual classifications. Where the FALSE row denotes where software's Face Bounding Boxes do not coincide with manually annotated faces. The tables shows that Google found 38.70% of the 90.34% of the Faces found that matched Faces also identified manually.

@@

```{r Image-Characteristics-Table, anchor="table"}
ImageCharacteristics <- ALLmetaIMG %>% filter(type=="Manual") %>% group_by(situation, bg, shotangle, detect) %>% dplyr::summarise(count=n())
ImageCharacteristics<-arrange(ImageCharacteristics, desc(count))
ImageCharacteristics<-ImageCharacteristics[1:10,]
kable(ImageCharacteristics, booktabs = TRUE)
```

`r figr("Image-Characteristics-Table", TRUE, type="Figure")`
 displays the feature combinations that produced the most potential face Bounding Boxes recognition by all four software. The most common shot is a crowd shot.
The second row in the table with 830 faces recognized is more interesting than the first result. This useful scene is an image of a Court Player Close-Up in front of a Logo Wall, taken at Player Shoulder Height.

@@

```{r Typical-Best-Recognised, results=FALSE, anchor="figure"}
imagesList1<-as.list("2016_CT6_R01_JMillman_AUS_vs_DSchwartzman_ARG_MS159_clip.0013.png") 
overlayGgplot(imagesList1, ALLmetaIMG, matchBox = FALSE)
```

```{r Google-Errors-Characteristics-Table, anchor="Table"}
library(dplyr)

@@

`r figr("Google-Errors-Characteristics-Table", TRUE, type="table")`
 shows that the potential Face Bounding Boxes Google returned were found in images that had the same characteristics of those that were manually annotated for faces.
<!-- This is shown by the same combinations of characteristics occuring in both Table 5 and Table 6.  keep as tables?-->
Without looking at each individual image this Table confirms that the potential Face Bounding Boxes Google located will likely be reasonable, and actually contain faces.

The best scenes for facial recognition have been found, given this information the following `r figr("Face-Characteristics-Table", TRUE, type="table")`
 considers the Characteristics of the individual faces found within those scenes. For this section we chose to consider only the faces that were manually annotated as players, with the intention of not basing results on recognition of undesirable faces.

```{r Face-Characteristics-Table, anchor="Table"}
FaceCharacteristics <- ALLmetaIMGPlayers %>%   filter(type=="Google")%>% filter(matchesManual==TRUE)%>% group_by(visorhat, glasses, headangle, lighting, obscured, detect) %>% dplyr::summarise(count=n())
FaceCharacteristics<-arrange(FaceCharacteristics, desc(count))
FaceCharacteristics<-as.tbl(FaceCharacteristics[1:10,])
kable(FaceCharacteristics, type = 'table')
```

`r figr("Face-Characteristics-Table", TRUE, type="table")`
utilizes a set of faces that were Manually annotated and also found by Google. 
Nine of the top ten facial characteristic combinations contained faces that were not wearing glasses.
The head angle describing the face angle was 'Other'^[Definition for head-angle factor level 'Other' found in..]
for nine of the top ten facial characteristic combinations.

```{r Glasses, anchor="Figure"}
GlassesGoogle <- ALLmetaIMGPlayers %>% filter(type=="Google") %>%
filter(matchesManual==TRUE)

pander(table(GlassesGoogle$glasses))
```

The extreme disparity between the amount of faces with glasses to without them means that the occurrence of these attributes across the software must be considered proportionally.

```{r Glasses-Figure, anchor="Figure"}
ggplotProportion(ALLmetaIMGPlayers, glasses)
```

Graph 2 above shows that Google outperforms the other software, with or without glasses. When there are no glasses worn Google finds over 60% of the manually annotated faces.

```{r HeadAngles, anchor="Table"}
HeadangleGoogle <- ALLmetaIMGPlayers %>%   filter(type=="Google") %>%
filter(matchesManual==TRUE)

@@

The amount of faces found by Google with the head-angle "Other", is much larger than the amount of faces with the head-angle "Profile" or "Front On". Therefore this should also be considered proportionately.

```{r Headangle-Figure, anchor="Figure"}
ggplotProportion(ALLmetaIMGPlayers, headangle)
```

Graph 3 displays that Google performs much better in comparison to the other software when the head-angle is Profile. This is outperforming unusually well, however this could also be due to the poor performance of the software in this circumstance.

@@

```{r Google-Characteristics-Table, anchor="table"}
GoogleCharacteristics <- ALLmetaIMG %>% filter(type=="Google") %>% filter(person=="Person") %>% filter(situation=="Court player close-up") %>% filter(bg=="Logo wall") %>% filter(detect=="Player") %>% group_by(person, situation, bg, shotangle, detect) %>% dplyr::summarise(count=n())
GoogleCharacteristics<-arrange(GoogleCharacteristics, desc(count))

@@

`r figr("Google-Characteristics-Table", TRUE, type="table")`
shows how many images with potential player's faces Google recognized. This displays a vast gap between the amount of potential faces found given the different shot angles. Shoulder height is an optimum angle for a Face Bounding Box.

>>>>

---
title: "Detecting Facial Expressions in Professional Tennis Matches"
author: "Stephanie Kobakian, Mitchell O'Hara-Wild, Dianne Cook, Stephanie Kovalcik"
date: '2017'
output: bookdown::pdf_document2
fontsize: 11pt
documentclass: article
bibliography: references.bib
---

```{r cache=FALSE, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
#packages:
library(bookdown)
library(pander)
library(ggplot2)
library(imager)
library(devtools)
library(knitcitations)
library(RefManageR)
library(readr)
library(knitr)
library(grid)
library(tidyverse)
library(magick)
library(kfigr)
library(UpSetR)
library(descr)
library(tables)

@@

#csv files:
ALLmetaIMG<-read_csv("ALLmetaIMG1.csv", col_types = cols(type = col_factor(levels = c("Manual", "Animetrics", "Google", "Microsoft", "Skybiometry"))))
SceneAttributes<-read_csv("TableCSVs/SceneAttributes.csv")
FaceAttributes<-read_csv("TableCSVs/FaceAttributes.csv")
SolutionSpecs<-read_csv("TableCSVs/SolutionSpecifications.csv")

@@

We were able to try the online demos to see whether these software were viable, we used the following Image displayed below in Figure \@ref(Trial-Image):

```{r Trial-Image, fig.cap = "Initial Image Used For Testing\\label{Trial-Image}", message=TRUE, warning=TRUE}
imagesList0<-as.list("2016_HSA_R01_BTomic_AUS_vs_DIstomin_UZB_MS157_clip.0013.png")
overlayGgplot(imagesList0, ALLmetaIMGFaces, matchBox = FALSE)
```

@@

```{r Scene-Attributes-Table, fig.cap ="\\label{tab:Scene-Attributes-Table}"}
kable(SceneAttributes, format = "pandoc", booktabs = TRUE, caption = "Attributes of a Scene of an image. The most appropriate option from the list was selected for each attribute.")
```

The aim was to collect specific information on each different face within the scene. To determine which of the sometimes many faces in the scene it would be reasonable for software to detect a standard was created for reasonable detection. This was based on the attributes provided in Table  \@ref(tab:Solution-Specifications-Table), below:

```{r Solution-Specifications-Table, results='asis', fig.cap = "\\label{tab:Solution-Specifications-Table}"}
kable(SolutionSpecs, format = "pandoc", booktabs = TRUE)

@@

```{r Face-Attribute-Tables, results='asis', fig.cap = "Attributes of an individual face within an image. The most appropriate option from the list was selected for each attribute.\\label{tab:Face-Attribute-Tables}"}
kable(FaceAttributes, format = "pandoc", booktabs = TRUE)
```

@@

The program we created was an interactive application that worked by presenting an image, using the imager @imager package, from the sample of images that had not yet been considered, appearing underneath were a set of radio buttons corresponding to the Scene Attributes in \@ref(tab:Scene-Attributes-Table) as shown above.

@@

Using the data set^[ALLmetaIMG] of the combined API and manual results, we were able to compare the performance of the software.
Firstly, we considered how many individual faces the software were able to detect in Figure \@ref(Face-Bounding-Box-Data).

However, individual faces are not beneficial if they do not correspond to the faces manually annotated. Figure \@ref(Face-Bounding-Box-Upset) was created by defining groups depending on the softwares that recognized each particular face. The UpSetR [-@UpSetR] package helps visualise set intersections. Where in this circumstance Face Bounding Boxes may overlap on the same face, each bar shows the number of individual faces that have Face Bounding Boxes resulting from each of the softwares highlighted below the bar.

@@

@@

```{r Face-Bounding-Box-Data, results='hide', fig.cap = "Face Bounding Boxes Per Software: The bar chart above shows the number of Bounding Boxes produced by each software, comparing the height of the bars indicates that Google's Facial Recognition software recognized almost 1000 more faces than the next best software, Microsoft.\\label{Face-Bounding-Box-Data}"}
ggplot(ALLmetaIMG, aes(x = type, fill=type)) + geom_bar(position="dodge") +xlab("Facial Recognition Softwares")+scale_x_discrete(limits=c("Manual","Google","Microsoft","Skybiometry","Animetrics")) +ylab("Number of Potential Faces") + guides(fill=FALSE)
```

@@

```{r Face-Bounding-Box-UpSet, fig.cap = "Face Bounding Boxes Per Software Combination. The Bar Chart shows the Face Bounding Boxes that were recognised by multiple software. The largest group with 809 faces is Face Bounding Boxes only found by Manual annotations. The following group was Boxes representing the same faces being recognized by both Manual annotations and the Google API.\\label{Face-Bounding-Box-Upset}"}
createUpSet(ALLmetaIMGFaces)
```

```{r I3, results=TRUE, fig.cap = "There is an unusual classification where just right of the center, the smaller box actually captures a fist, not a face.\\label{fig:I3}"}
imagesList3<-as.list("2016_SC2_R01_ATomljanovic_AUS_vs_KBondarenko_UKR_WS1112_clip.0053.png") 
overlayGgplot(imagesList3, ALLmetaIMGFaces, matchBox = FALSE)
```

@@

```{r Type-Matches-Manual, fig.cap = "\\label{tab:Type-Matches-Manual}"}

NoManual<-ALLmetaIMG %>% filter(ALLmetaIMG$type!="Manual")
NoManual<-NoManual %>% droplevels(NoManual$type) 
Accuracy<-with(NoManual, table(NoManual$matchesManual, NoManual$type))
CrossTable(Accuracy, prop.chisq = FALSE, cell.layout = TRUE, digits=4)
```

Table \@ref(Type-Matches-Manual) shows whether the potential Face Bounding Boxes match faces that were annotated during the manual classifications. Where the FALSE row denotes where software's Face Bounding Boxes do not coincide with manually annotated faces. The tables shows that Google found 38.70% of the 90.34% of the Faces found that matched Faces also identified manually.

@@

```{r Image-Characteristics-Table, fig.cap = "\\label{tab:Image-Characteristics-Table}"}
ImageCharacteristics <- ALLmetaIMG %>% filter(type=="Manual") %>% group_by(situation, bg, shotangle, detect) %>% dplyr::summarise(count=n())
ImageCharacteristics<-arrange(ImageCharacteristics, desc(count))
ImageCharacteristics<-ImageCharacteristics[1:10,]
kable(ImageCharacteristics, booktabs = TRUE)
```

\@ref(Image-Characteristics-Table) displays the feature combinations that produced the most potential face Bounding Boxes recognition by all four software. The most common shot is a crowd shot.
The second row in the table with 830 faces recognized is more interesting than the first result. This useful scene is an image of a Court Player Close-Up in front of a Logo Wall, taken at Player Shoulder Height.

@@

```{r Typical-Best-Recognised, results=FALSE, fig.cap = "\\label{fig:Typical-Best-Recognised}"}
imagesList1<-as.list("2016_CT6_R01_JMillman_AUS_vs_DSchwartzman_ARG_MS159_clip.0013.png")
overlayGgplot(imagesList1, ALLmetaIMG, matchBox = FALSE)
```

```{r Google-Errors-Characteristics-Table, fig.cap = "\\label{tab:Google-Errors-Characteristics-Table}"}
library(dplyr)

@@

Table \@ref(Google-Errors-Characteristics-Table)
 shows that the potential Face Bounding Boxes Google returned were found in images that had the same characteristics of those that were manually annotated for faces.
<!-- This is shown by the same combinations of characteristics occuring in both Table 5 and Table 6.  keep as tables?-->
Without looking at each individual image this Table confirms that the potential Face Bounding Boxes Google located will likely be reasonable, and actually contain faces.

The best scenes for facial recognition have been found, given this information the following \@ref(Face-Characteristics-Table) considers the Characteristics of the individual faces found within those scenes. For this section we chose to consider only the faces that were manually annotated as players, with the intention of not basing results on recognition of undesirable faces.

```{r Face-Characteristics-Table, fig.cap = "\\label{tab:Face-Characteristics-Table}"}
FaceCharacteristics <- ALLmetaIMGPlayers %>%   filter(type=="Google")%>% filter(matchesManual==TRUE)%>% group_by(visorhat, glasses, headangle, lighting, obscured, detect) %>% dplyr::summarise(count=n())
FaceCharacteristics<-arrange(FaceCharacteristics, desc(count))
FaceCharacteristics<-as.tbl(FaceCharacteristics[1:10,])
kable(FaceCharacteristics, type = 'table')
```

Table \@ref(Face-Characteristics-Table)
utilizes a set of faces that were Manually annotated and also found by Google. 
Nine of the top ten facial characteristic combinations contained faces that were not wearing glasses.
The head angle describing the face angle was 'Other'^[Definition for head-angle factor level 'Other' found in..]
for nine of the top ten facial characteristic combinations.

```{r Glasses, fig.cap = "This table shows the amount of faces recognised by google wearing glasses. The strong disparity between the amount of faces with glasses to faces without them means that the occurrence of these attributes across the software must be considered proportionally.\\label{tab:Glasses}"}
GlassesGoogle <- ALLmetaIMGFaces %>% filter(type=="Google") %>%
filter(matchesManual==TRUE)
pander(table(GlassesGoogle$glasses))
```

```{r Glasses-Figure, fig.cap = "The figure above shows that Google outperforms the other software, with or without glasses. When there are no glasses worn Google finds over 60% of the manually annotated faces.\\label{fig:Glasses-Figure}"}
ggplotProportion(ALLmetaIMGPlayers, glasses)+xlab("Glasses on Face")
```

```{r HeadAngles, fig.cap = "\\label{tab:HeadAngles}"}
HeadangleGoogle <- ALLmetaIMGPlayers %>%   filter(type=="Google") %>%
filter(matchesManual==TRUE)

@@

In Table \@ref(HeadAngles), the amount of faces found by Google with the head-angle "Other", is much larger than the amount of faces with the head-angle "Profile" or "Front On". Therefore this should also be considered proportionately.

```{r Headangle-Figure, fig.cap = "Google performs much better in comparison to the other software when the head-angle is Profile. This is outperforming unusually well, however this could also be due to the poor performance of the software in this circumstance.\\label{fig:Headangle-Figure}"}
ggplotProportion(ALLmetaIMGPlayers, headangle)+xlab("Head Angle of Face")
```

@@

```{r Google-Characteristics-Table, fig.cap = "\\label{tab:Google-Characteristics-Table}"}
GoogleCharacteristics <- ALLmetaIMG %>% filter(type=="Google") %>% filter(person=="Person") %>% filter(situation=="Court player close-up") %>% filter(bg=="Logo wall") %>% filter(detect=="Player") %>% group_by(person, situation, bg, shotangle, detect) %>% dplyr::summarise(count=n())
GoogleCharacteristics<-arrange(GoogleCharacteristics, desc(count))

@@

Table \@ref(Google-Characteristics-Table)
shows how many images with potential player's faces Google recognized. This displays a vast gap between the amount of potential faces found given the different shot angles. Shoulder height is an optimum angle for a Face Bounding Box.
